---
title: "Data preprocessing"
output: html_document
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, hms)
```

# 1) Data overview
  A) 'Audio' contains recordings from the interviews in wav format.
  B) 'Diarization' contains turn-taking data automatically extracted from the audio recordings. There are raw data in .seg format and cleaned up data in .csv.
  C) 'Gesture' containts raw data from the actigraphs (wrist movement sensors). This was originally measured as changes in acceleration on 3 axis which were meaningless and so euclidian distance was calculated.
  D) 'Timing_Handedness.txt' contains hand-notes about start time of the gesturing and voice start, handedness of the subject and schizophrenia symptoms.

## 1.1) Look at one sample from all data types
```{r first look}
#turn-taking sample
turn = read.csv("raw_data/Diarization/101.csv")

#gesture sample
gest = read.csv('raw_data/Gesture/Sub101RightHanded.csv')
```

```{r clean Timing_Handedness.txt}
#strinAsFactors=F because time-data would get converted to factors
Timing = read.delim("raw_data/Timing_Handedness.txt", na.strings="NaN", stringsAsFactors=F)

#all time-data needs to be converted to hms format from strings - I'll need to calculate with them
#Diff is difference between VoiceStart and GestureStart in seconds
Timing = Timing %>% mutate(
  GestureStart=as.hms(GestureStart),
  VoiceStart=as.hms(VoiceStart),
  Diff=VoiceStart-GestureStart
)
```

  
